Review of Edge Detectors

### 2014-11-11 old_versions/edge-detection-review-v2.tex
second round of third revision, improving references

\usepackage{hyperref,verbatim,graphicx,amsmath,amssymb,amssymb,dsfont}
\usepackage[ruled,linesnumbered]{algorithm2e}
\usepackage{sidecap,subfig,array}
\usepackage[table]{xcolor}
\usepackage{multicol}
\newtheorem{theorem}{Theorem}
\numberwithin{equation}{section}
\numberwithin{table}{section}
\numberwithin{figure}{section}
\begin{document}

@@

%-------------------------------------------------------------------------------
\begin{ipolAbstract}
In this paper we study some of the classic alternatives for edge detection in digital images. The main idea 
behind the edge detection is to find where there are abrupt changes in the intensity of an image. 
As a first approach, the first derivative is used to find the changes of intensity in several algorithms 
such as Sobel, Prewitt and Roberts. The second derivative is also used, for example in algorithms 
like Marr-Hildreth and Haralick.\\
We compare the results obtained from a qualitative point of view (perceptual) and from a quantitative 
point of view (number of operations, execution time), considering different forms to convolve an 
image with a kernel (step required in some of the algorithms).\\

@@

Edge detection is one of the first approaches published in the literature for 
segmenting images. The basic idea is to detect abrupt changes in intensity. 
Detecting those changes in intensity for the purpose of finding edges in images 
can be accomplished using first or second order derivatives.

In the 70's, edge detection methods were based on using small operators 
(such as Sobel masks), attempting to compute an approximation of the
first derivative of the image. In section ~\ref{sec:first} we study the 
performance of such algorithms, as an introduction to a more sophisticated 
analysis into the edge finding process.\\

@@

\begin{SCfigure}[3][!h]
	\centering
	\includegraphics[width=0.14\textwidth]{1dmasks.pdf}
	\caption{One-dimensional masks used to implement equations \ref{eq:roberts1} and \ref{eq:roberts2}.}
	\label{fig:1dmasks}
\end{SCfigure}

\begin{SCfigure}[3][!h]
	\centering
	\includegraphics[width=0.15\textwidth]{roberts.pdf}
	\caption{\textit{Roberts cross-gradient} 2-D masks.}
	\label{fig:roberts}
\end{SCfigure}

@@

\begin{SCfigure}[2][!h]
	\centering
	\includegraphics[width=0.25\textwidth]{prewitt.pdf}
	\caption{\textit{Prewitt} 2-D masks of size $3\times3$.}
	\label{fig:prewitt}
\end{SCfigure}

@@

\begin{SCfigure}[2][!h]
	\centering
	\includegraphics[width=0.25\textwidth]{sobel.pdf}
	\caption{\textit{Sobel} 2-D masks of size $3\times3$.}
	\label{fig:sobel}
\end{SCfigure}

@@

@@

First we see the results of each of the algorithms on a simple test image, the white square on black background shown in figure \ref{fig:original1}. \\

\begin{figure}[h!]
	\centering
	\includegraphics[width=0.2\textwidth]{results/square.png}
	\caption{Test image 1: white square on black background.}
	\label{fig:original1}
\end{figure}

The figures \ref{fig:result1-a} to \ref{fig:result1-f} show the output of each algorithm. The first three correspond to the first derivative methods (Roberts, Prewitt and Sobel), then the following two are from the Marr-Hildreth algorithm using Gaussian and LoG kernels, and the last one corresponds to the Haralick algorithm. Execution time\footnote{Running on Intel Core i3 CPU (2.53GHz), 3 Gb RAM, standard laptop.} for each algorithm are shown in Table \ref{exectime1}. \\

\begin{figure}[h!]
	\centering
	\subfloat[Roberts.]{\label{fig:result1-a}\hspace*{0.5cm}\includegraphics[width=0.2\textwidth]{results/square_roberts.png}\hspace*{0.5cm}}
	\quad
	\subfloat[Prewitt.]{\label{fig:result1-b}\hspace*{0.5cm}\includegraphics[width=0.2\textwidth]{results/square_prewitt.png}\hspace*{0.5cm}}
	\quad
	\subfloat[Sobel.]{\label{fig:result1-c}\hspace*{0.5cm}\includegraphics[width=0.2\textwidth]{results/square_sobel.png}\hspace*{0.5cm}}

	\subfloat[Marr-Hildreth (Gaussian).]{\label{fig:result1-d}\hspace*{0.5cm}\includegraphics[width=0.2\textwidth]{results/square_marr-hildreth.png}\hspace*{0.5cm}}
	\quad
	\subfloat[Marr-Hildreth (LoG).]{\label{fig:result1-e}\hspace*{0.5cm}\includegraphics[width=0.2\textwidth]{results/square_marr-hildreth-log.png}\hspace*{0.5cm}}
	\quad
	\subfloat[Haralick.]{\label{fig:result1-f}\hspace*{0.5cm}\includegraphics[width=0.2\textwidth]{results/square_haralick.png}\hspace*{0.5cm}}
	\caption{Results of the algorithms using the figure \ref{fig:original1} as input image.}
	\label{fig:result1}
\end{figure}

\begin{table}[t!]
	\begin{center}
	\begin{tabular}{| l | r |}
		\hline \rule{0pt}{3ex}
		\cellcolor[gray]{0.8} \textbf{Algorithm}	& \cellcolor[gray]{0.8} \textbf{Execution time (s)}	\\ \hline \rule{0pt}{3ex}
		Roberts, Prewitt and Sobel					& $0.020 \ s$										\\ \hline \rule{0pt}{3ex}
		Marr-Hildreth (Gaussian)					& $0.030 \ s$										\\ \hline \rule{0pt}{3ex}
		Marr-Hildreth (LoG)							& $0.050 \ s$										\\ \hline \rule{0pt}{3ex}
		Haralick									& $0.040 \ s$										\\
		\hline
	\end{tabular}
	\end{center}
	\caption{Execution time of the algorithms using the figure \ref{fig:original1} as input image (128$\times$128 pixels, 1 channel).}
	\label{exectime1}
\end{table}
\vspace{0.5cm}

It is appreciated that in the simplest cases, the first derivative edge detection algorithms are running better and faster. Let us now analyze the performance of each of the algorithmos using a more realistic test image, shown in figure \ref{fig:original2}. \\

\begin{figure}[t!]
	\centering
	\includegraphics[width=0.5\textwidth]{results/molino.jpg}
	\caption{Test image 2: windmill.}
	\label{fig:original2}
\end{figure}

First we show in figure \ref{fig:result2} the results of the first derivative edge detection algorithms.

\begin{figure}[h!]
	\centering
	\subfloat[Roberts.]{\label{fig:result2-a}\includegraphics[width=0.42\textwidth]{results/molino_roberts.png}}
	\quad
	\subfloat[Prewitt.]{\label{fig:result2-b}\includegraphics[width=0.42\textwidth]{results/molino_prewitt.png}}

	\subfloat[Sobel.]{\label{fig:result2-c}\includegraphics[width=0.42\textwidth]{results/molino_sobel.png}}
	\caption{Results of the first derivative algorithms using the figure \ref{fig:original2} as input image.}
	\label{fig:result2}
\end{figure}

Now we show in figures \ref{fig:result3-a} and \ref{fig:result3-b} the results obtained using the Marr-Hildreth algorithm (both Gaussian and LoG kernels), and in figure \ref{fig:result3-c} the results obtained using the Haralick algorithm. In table \ref{exectime2} we show the execution time for each one of the algorithms. \\

\begin{figure}[h!]
	\centering
	\subfloat[Marr-Hildreth (Gaussian).]{\label{fig:result3-a}\includegraphics[width=0.42\textwidth]{results/molino_marr-hildreth.png}}
	\quad
	\subfloat[Marr-Hildreth (LoG).]{\label{fig:result3-b}\includegraphics[width=0.42\textwidth]{results/molino_marr-hildreth-log.png}}

	\subfloat[Haralick.]{\label{fig:result3-c}\includegraphics[width=0.42\textwidth]{results/molino_haralick.png}}
	\caption{Results of the Marr-Hildreth and the Haralick algorithms, using the figure \ref{fig:original2} as input image.}
	\label{fig:result3}
\end{figure}

\begin{table}[h!]
	\begin{center}
	\begin{tabular}{| l | r |}
		\hline \rule{0pt}{3ex}
		\cellcolor[gray]{0.8} \textbf{Algorithm}	& \cellcolor[gray]{0.8} \textbf{Execution time (s)}	\\ \hline \rule{0pt}{3ex}
		Roberts, Prewitt and Sobel					& $0.550 \ s$										\\ \hline \rule{0pt}{3ex}
		Marr-Hildreth (Gaussian)					& $1.050 \ s$										\\ \hline \rule{0pt}{3ex}
		Marr-Hildreth (LoG)							& $1.440 \ s$										\\ \hline \rule{0pt}{3ex}
		Haralick									& $0.930 \ s$										\\
		\hline
	\end{tabular}
	\end{center}
	\caption{Execution time of the algorithms using the figure \ref{fig:original2} as input image (1000$\times$563 pixels, 3 channels).}
	\label{exectime2}
\end{table}
\vspace{0.5cm}

It is now visible the difference in performance between the first derivative edge detection algorithms and the second derivative ones. \\

\clearpage
%-------------------------------------------------------------------------------
\section{Conclusions}
\label{sec:conclusions}

\clearpage
%-------------------------------------------------------------------------------
\section{Appendix: Common operations}
\label{sec:appendix1}

@@

\begin{SCfigure}[][!t]
	\centering
	\includegraphics[width=0.5\textwidth]{kernel_gaussian.pdf}
	\caption{Gaussian kernel, $\sigma=4$, $n=25$. Is easy to see that the selected value of n is 
sufficient to have a good approximation of the Gaussian function in the kernel.}
	\label{fig:gaussian_kernel}
\end{SCfigure}

@@

\begin{SCfigure}[][!t]
	\centering
	\includegraphics[width=0.5\textwidth]{kernel_log.pdf}
	\caption{Laplacian of a Gaussian kernel, $\sigma=4$, $n=31$. The selected value of n is sufficient 
to have a good approximation of the LoG function in the kernel, but is greater than in the case of 
Gaussian kernel.}
	\label{fig:log_kernel}
\end{SCfigure}

@@

\begin{SCfigure}[][!t]
	\centering
	\includegraphics[width=0.5\textwidth]{kernels.pdf}
	\caption{Comparison of the Gaussian and LoG functions.}
	\label{fig:kernels}
\end{SCfigure}

\clearpage
%-------------------------------------------------------------------------------
\section{Examples}
\label{sec:examples}

\subsection{Images}

Further examples obtained with the implemented algorithms are shown in figures \ref{fig:example1} to \ref{fig:example3}. More examples can be found in the online demo. \\

\begin{figure}[h!]
	\centering
	\subfloat[Roberts.]{\includegraphics[width=0.3\textwidth]{examples/lena.png}}

	\subfloat[Roberts.]{\includegraphics[width=0.3\textwidth]{examples/lena_roberts.png}}
	\quad
	\subfloat[Prewitt.]{\includegraphics[width=0.3\textwidth]{examples/lena_prewitt.png}}
	\quad
	\subfloat[Sobel.]{\includegraphics[width=0.3\textwidth]{examples/lena_sobel.png}}

	\subfloat[Marr-Hildreth (Gaussian).]{\includegraphics[width=0.3\textwidth]{examples/lena_marr-hildreth.png}}
	\quad
	\subfloat[Marr-Hildreth (LoG).]{\includegraphics[width=0.3\textwidth]{examples/lena_marr-hildreth-log.png}}
	\quad
	\subfloat[Haralick.]{\includegraphics[width=0.3\textwidth]{examples/lena_haralick.png}}
	\caption{Example: Lena.}
	\label{fig:example1}
\end{figure}

\begin{figure}[t!]
	\centering
	\subfloat[Roberts.]{\includegraphics[width=0.3\textwidth]{examples/oranges.png}}

	\subfloat[Roberts.]{\includegraphics[width=0.3\textwidth]{examples/oranges_roberts.png}}
	\quad
	\subfloat[Prewitt.]{\includegraphics[width=0.3\textwidth]{examples/oranges_prewitt.png}}
	\quad
	\subfloat[Sobel.]{\includegraphics[width=0.3\textwidth]{examples/oranges_sobel.png}}

	\subfloat[Marr-Hildreth (Gaussian).]{\includegraphics[width=0.3\textwidth]{examples/oranges_marr-hildreth.png}}
	\quad
	\subfloat[Marr-Hildreth (LoG).]{\includegraphics[width=0.3\textwidth]{examples/oranges_marr-hildreth-log.png}}
	\quad
	\subfloat[Haralick.]{\includegraphics[width=0.3\textwidth]{examples/oranges_haralick.png}}
	\caption{Example: Oranges.}
	\label{fig:example2}
\end{figure}

\vspace{2cm}

\begin{figure}[t!]
	\centering
	\subfloat[Roberts.]{\includegraphics[width=0.3\textwidth]{examples/shapes.png}}

	\subfloat[Roberts.]{\includegraphics[width=0.3\textwidth]{examples/shapes_roberts.png}}
	\quad
	\subfloat[Prewitt.]{\includegraphics[width=0.3\textwidth]{examples/shapes_prewitt.png}}
	\quad
	\subfloat[Sobel.]{\includegraphics[width=0.3\textwidth]{examples/shapes_sobel.png}}

	\subfloat[Marr-Hildreth (Gaussian).]{\includegraphics[width=0.3\textwidth]{examples/shapes_marr-hildreth.png}}
	\quad
	\subfloat[Marr-Hildreth (LoG).]{\includegraphics[width=0.3\textwidth]{examples/shapes_marr-hildreth-log.png}}
	\quad
	\subfloat[Haralick.]{\includegraphics[width=0.3\textwidth]{examples/shapes_haralick.png}}
	\caption{Example: Shapes.}
	\label{fig:example3}
\end{figure}

@@

%TODO: videos de ejemplo con todos los métodos.

This is an example of applying the edge detection algorithms described here, frame by frame, to a video: 
\begin{itemize}
	\centering
	\item \href{http://iie.fing.edu.uy/~haldos/ipol/video.mov}{original} (xxxMb).
\end{itemize}
\begin{multicols}{2}
\begin{itemize}
	\item \href{http://iie.fing.edu.uy/~haldos/ipol/video-roberts.mov}{roberts version} (xxxMb).
	\item \href{http://iie.fing.edu.uy/~haldos/ipol/video-prewitt.mov}{prewitt version} (xxxMb).
	\item \href{http://iie.fing.edu.uy/~haldos/ipol/video-sobel.mov}{sobel version} (xxxMb).
	\item \href{http://iie.fing.edu.uy/~haldos/ipol/video-marr-hildreth-gaussian.mov}{marr-hildreth-gaussian version} (xxxMb).
	\item \href{http://iie.fing.edu.uy/~haldos/ipol/video-marr-hildreth-log.mov}{marr-hildreth-log version} (xxxMb).
	\item \href{http://iie.fing.edu.uy/~haldos/ipol/video-haralick.mov}{haralick version} (xxxMb).
\end{itemize}
\end{multicols}

>>>>

\usepackage{hyperref,verbatim,graphicx,amsmath,amssymb,amssymb,dsfont}
\usepackage[ruled,linesnumbered]{algorithm2e}
\usepackage{sidecap,subfig,array}
\newtheorem{theorem}{Theorem}
\numberwithin{equation}{section}
\numberwithin{table}{section}
\begin{document}

@@

%-------------------------------------------------------------------------------
\begin{ipolAbstract}
In this paper we study different alternatives for edge detection in digital images. The main idea 
behind the edge detection is to find where there are abrupt changes in the intensity of an image. 
As a first approach, the first derivative is used to find the changes of intensity in several algorithms 
such as Sobel, Prewitt and Roberts. The second derivative is also used, for example in algorithms 
like Marr-Hildreth and Haralick.\\
We compare the results obtained from a qualitative point of view (perceptual) and from a quantitative 
point of view (number of operations, execution time), considering different forms to convolve an 
image with a kernel (step required in some of the algorithms).\\

@@

Edge detection is the first approach for segmenting images
based on abrupt changes in intensity. Detecting those changes in intensity 
for the purpose of finding edges in images can be accomplished using first 
or second order derivatives.

In the 70's, edge detection methods were based on using small operators 
(such as Sobel masks), attempting to compute an approximation of the
first derivative of the image. In section ~\ref{sec:first} we study the 
performance of these algorithms, as an introduction to a more sophisticated 
analysis into the edge finding process.\\

@@

\begin{SCfigure}[3][!h]
	\centering
	\includegraphics[width=0.14\textwidth]{1dmasks.eps}
	\caption{One-dimensional masks used to implement equations \ref{eq:roberts1} and \ref{eq:roberts2}.}
	\label{fig:1dmasks}
\end{SCfigure}

\begin{SCfigure}[3][!h]
	\centering
	\includegraphics[width=0.15\textwidth]{roberts.eps}
	\caption{\textit{Roberts cross-gradient} 2-D masks.}
	\label{fig:roberts}
\end{SCfigure}

@@

\begin{SCfigure}[2][!h]
	\centering
	\includegraphics[width=0.25\textwidth]{prewitt.eps}
	\caption{\textit{Prewitt} 2-D masks of size $3\times3$.}
	\label{fig:prewitt}
\end{SCfigure}

@@

\begin{SCfigure}[2][!h]
	\centering
	\includegraphics[width=0.25\textwidth]{sobel.eps}
	\caption{\textit{Sobel} 2-D masks of size $3\times3$.}
	\label{fig:sobel}
\end{SCfigure}

@@

%\begin{align}
%	\label{eq:haralick_masks}
%	\mathbf{b_1} & = 	\begin{bmatrix} 425 & 275 & 225 & 275 & 425 \\ 
%										275 & 125 &  75 & 125 & 275 \\ 
%										225 &  75 &  25 &  75 & 225 \\
%										275 & 125 &  75 & 125 & 275 \\
%										425 & 275 & 225 & 275 & 425 
%						\end{bmatrix} &
%	\mathbf{b_2} & =	\begin{bmatrix} -2260 & -620 & 0 & 620 & 2260 \\ 
%										-1660 & -320 & 0 & 320 & 1660 \\ 
%										-1460 & -220 & 0 & 220 & 1460 \\
%										-1660 & -320 & 0 & 320 & 1660 \\ 
%										-2260 & -620 & 0 & 620 & 2260 
%						\end{bmatrix} \nonumber \\
%	\mathbf{b_3} & =	\begin{bmatrix} 2260  &  1660 &  1460 &  1660 &  2260 \\ 
%									  	620   &   320 &   220 &   320 &   620 \\ 
%										0     &     0 &     0 &     0 &     0 \\
%										-620  &  -320 &  -220 &  -320 &  -620 \\ 
%										-2260 & -1660 & -1460 & -1660 & -2260
%						\end{bmatrix} &
%	\mathbf{b_4} & =	\begin{bmatrix} 1130 & 620 & 450 & 620 & 1130 \\ 
%									  	830  & 320 & 150 & 320 &  830 \\ 
%										730  & 220 &  50 & 220 &  730 \\
%										830  & 320 & 150 & 320 &  830 \\ 
%										1130 & 620 & 450 & 620 & 1130
%						\end{bmatrix} \nonumber \\
%	\mathbf{b_5} & =	\begin{bmatrix} -400 & -200 & 0 &  200 &  400 \\ 
%									  	-200 & -100 & 0 &  100 &  200 \\ 
%										0    &    0 & 0 &    0 &    0 \\
%										200  &  100 & 0 & -100 & -200 \\ 
%										400  &  200 & 0 & -200 & -400
%						\end{bmatrix} &
%	\mathbf{b_6} & =	\begin{bmatrix} 1130 & 830 & 730 & 830 & 1130 \\ 
%									  	620  & 320 & 220 & 320 &  620 \\ 
%										450  & 150 &  50 & 150 &  450 \\
%										620  & 320 & 220 & 320 &  620 \\ 
%										1130 & 830 & 730 & 830 & 1130
%						\end{bmatrix} \nonumber \\
%	\mathbf{b_7} & =	\begin{bmatrix} -8260 & -2180 & 0 & 2180 & 8260 \\ 
%									  	-6220 & -1160 & 0 & 1160 & 6220 \\ 
%										-5540 &  -820 & 0 &  820 & 5540 \\
%										-6220 & -1160 & 0 & 1160 & 6220 \\ 
%										-8260 & -2180 & 0 & 2180 & 8260
%						\end{bmatrix} &
%	\mathbf{b_8} & =	\begin{bmatrix} 5640  &  3600 &  2920 &  3600 &  5640 \\ 
%									  	1800  &   780 &   440 &   780 &  1800 \\ 
%										0     &     0 &     0 &     0 &     0 \\
%										-1800 &  -780 &  -440 &  -780 & -1800 \\ 
%										-5640 & -3600 & -2920 & -3600 & -5640
%						\end{bmatrix} \nonumber \\
%	\mathbf{b_9} & =	\begin{bmatrix} -5640 & -1800 & 0 & 1800 & 5640 \\ 
%									  	-3600 &  -780 & 0 &  780 & 3600 \\ 
%										-2920 &  -440 & 0 &  440 & 2920 \\
%										-3600 &  -780 & 0 &  780 & 3600 \\ 
%										-5640 & -1800 & 0 & 1800 & 5640
%						\end{bmatrix} &
%	\mathbf{b_{10}} & =	\begin{bmatrix} 8260  &  6220 &  5540 &  6220 &  8260 \\ 
%									  	2180  &  1160 &   820 &  1160 &  2180 \\ 
%										0     &     0 &     0 &     0 &     0 \\
%										-2180 & -1160 &  -820 & -1160 & -2180 \\ 
%										-8260 & -6220 & -5540 & -6220 & -8260
%						\end{bmatrix} \nonumber \\
%\end{align}

@@

Let us analyze the performance of each of the methods described above using some test images. The first image 
is a simple example, a white square on black background. In figure \ref{fig:result1} we can see the results 
of applying each of the algorithms to this image.

%-------------------------------------------------------------------------------
\section{Conclusions}
\label{sec:conclusions}

%-------------------------------------------------------------------------------
\section{Appendix: Common operations}
\label{sec:appendix1}

@@

\begin{SCfigure}[][!t]
	\centering
	\includegraphics[width=0.5\textwidth]{kernel_gaussian.eps}
	\caption{Gaussian kernel, $\sigma=4$, $n=25$. Is easy to see that the selected value of n is 
sufficient to have a good approximation of the Gaussian function in the kernel.}
	\label{fig:gaussian_kernel}
\end{SCfigure}

@@

\begin{SCfigure}[][!t]
	\centering
	\includegraphics[width=0.5\textwidth]{kernel_log.eps}
	\caption{Laplacian of a Gaussian kernel, $\sigma=4$, $n=31$. The selected value of n is sufficient 
to have a good approximation of the LoG function in the kernel, but is greater than in the case of 
Gaussian kernel.}
	\label{fig:log_kernel}
\end{SCfigure}

@@

\begin{SCfigure}[][!b]
	\centering
	\includegraphics[width=0.5\textwidth]{kernels.eps}
	\caption{Comparison of the Gaussian and LoG functions.}
	\label{fig:kernels}
\end{SCfigure}

%-------------------------------------------------------------------------------
\section{Examples}
\label{sec:examples}

@@

This is an example of applying the Marr-Hildreth edge detector, frame by frame, to a video:
\begin{itemize}
	\item \href{http://iie.fing.edu.uy/~haldos/ipol/video.mov}{original}(xxxMb).
	\item \href{http://iie.fing.edu.uy/~haldos/ipol/video-marr-hildreth.mov}{marr-hildreth version}(xxxMb).
\end{itemize}
